---
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    includes:
      #in_header: ../../../../../_Fns/packages4looks4rmd_beamer.tex
#title: 'Supply and Demand'
#author: "Abel Embaye"
#date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r , include=FALSE}
#knitr::knit_exit()
library(psych)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(cowplot)
#install.packages("ggpmisc")                         # Install & load ggpmisc package
library("ggpmisc")
#install.packages("patchwork")                       # Install & load patchwork
library("patchwork")
library(ggpmisc)
library(gridExtra)
library(grid)
library("ggdensity")
library(ggpubr)
library(tidyverse)
library(colorspace) # for darken()
library(ggrepel)    # for geom_text_repel()
```


```{r , echo=FALSE, message=F, warning=F, fig.align='center'}
xvar_name= "Gasoline price (in dollars per gallon)"
yvar_name="Gasoline consumption\n (gallon per day per person)"
p<- c(3, 4, 5.5,6, 7.5,7.6, 7.8, 8)
country<- c("U.S.", "Canada", "Japan", "Spain",
      "Italy", "France", "U.K.", "Germany")
region<- c("Americas", "Americas", "Asia", "Europe",
      "Europe", "Europe", "Europe", "Europe")
q<- c(1.3, 1.0, .4, .2,.35, .2,.38, .37 )

df<- data.frame(q,p, country, region)
nregions= length(unique(df$region))
#"country" "region"  "cpi"     "hdi"

# Okabe Ito colors

# region_cols <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2" , "#999999")
# Define a color palette
color_palette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#999999", "#D55E00", "#CC79A7", "#F0E442", "#0072B2")

region_cols <-  color_palette[1:(nregions+1)]
country_highlight <- c("U.S.", "Canada", "Japan", "Spain",
      "Italy", "France", "U.K.", "Germany")

df <- df %>%
  mutate(
    label = ifelse(country %in% country_highlight, country, "")
  )

df <- df %>%
  mutate(region = case_when(
    region == "Middle East and North Africa" ~ "Middle East\nand North Africa",
    region == "Europe and Central Asia" ~ "Europe and\nCentral Asia",
    region == "Sub Saharan Africa" ~ "Sub-Saharan\nAfrica",
    TRUE ~ region)
    )

ggplot(df, aes(q, p)) +
  geom_smooth(
    aes(color = "y ~ log(x)", fill = "y ~ log(x)"),
    method = 'lm', formula = y~log(x), se = FALSE, fullrange = TRUE
  ) +
  geom_point(
    aes(color = region, fill = region),
    size = 2.5, alpha = 0.5, shape = 21
  ) +
  geom_text_repel(
    aes(label = label), # this label are the selected countries
    color = "black",
    size = 9/.pt, # font size 9 pt
    point.padding = 0.1, 
    box.padding = .6,
    min.segment.length = 0,
    max.overlaps = 1000,
    seed = 7654
  ) +
  scale_color_manual(
    name = NULL,
    values = darken(region_cols, 0.3)
  ) +
  scale_fill_manual(
    name = NULL,
    values = region_cols
  ) +
  scale_x_continuous(
    name = xvar_name ,
    limits = c(min(q)-.5, max(q)+.5),
    breaks = c(0, .25, .50, 1.0, 1.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = yvar_name,
    limits = c(min(p)-1, max(p)),
    breaks = 2:8,
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      nrow = 1,
      override.aes = list(
        linetype = c(rep(0, nregions), 1),
        shape = c(rep(21, nregions), NA)
      )
    )
  ) +
  theme_minimal_hgrid(12, rel_small = 1) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt")
  )

```

---

```{r , echo=FALSE, message=F, warning=F, fig.align='center'}
load(file="C:/Users/aembaye/OneDrive - University of Arkansas/C2-embaye/MyProjects_new/corruption.rda")
#View(corruption)
df <- corruption %>% 
  filter(year == 2015) %>% 
  na.omit()

# Okabe Ito colors
region_cols <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2" , "#999999")

country_highlight <- c("Germany", "Norway", "United States", "Greece", "Singapore", "Rwanda", "Russia", "Venezuela", "Sudan", "Iraq", "Ghana", "Niger", "Chad", "Kuwait", "Qatar", "Myanmar", "Nepal", "Chile", "Argentina", "Japan", "China")

df <- df %>%
  mutate(
    label = ifelse(country %in% country_highlight, country, "")
  )

df <- df %>%
  mutate(region = case_when(
    region == "Middle East and North Africa" ~ "Middle East\nand North Africa",
    region == "Europe and Central Asia" ~ "Europe and\nCentral Asia",
    region == "Sub Saharan Africa" ~ "Sub-Saharan\nAfrica",
    TRUE ~ region)
    )

ggplot(df, aes(cpi, hdi)) +
  geom_smooth(
    aes(color = "y ~ log(x)", fill = "y ~ log(x)"),
    method = 'lm', formula = y~log(x), se = FALSE, fullrange = TRUE
  ) +
  geom_point(
    aes(color = region, fill = region),
    size = 2.5, alpha = 0.5, shape = 21
  ) +
  geom_text_repel(
    aes(label = label),
    color = "black",
    size = 9/.pt, # font size 9 pt
    point.padding = 0.1, 
    box.padding = .6,
    min.segment.length = 0,
    max.overlaps = 1000,
    seed = 7654
  ) +
  scale_color_manual(
    name = NULL,
    values = darken(region_cols, 0.3)
  ) +
  scale_fill_manual(
    name = NULL,
    values = region_cols
  ) +
  scale_x_continuous(
    name = "Corruption Perceptions Index, 2015 (100 = least corrupt)",
    limits = c(10, 95),
    breaks = c(20, 40, 60, 80, 100),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = "Human Development Index, 2015\n(1.0 = most developed)",
    limits = c(0.3, 1.05),
    breaks = c(0.2, 0.4, 0.6, 0.8, 1.0),
    expand = c(0, 0)
  ) +
  guides(
    color = guide_legend(
      nrow = 1,
      override.aes = list(
        linetype = c(rep(0, 5), 1),
        shape = c(rep(21, 5), NA)
      )
    )
  ) +
  theme_minimal_hgrid(12, rel_small = 1) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt")
  )

```

---
```{r , echo=FALSE, message=F, warning=F, fig.width=15}
# Create the data frame
p = 1:10
q = 2 * (1:10)
draw_and_table <- function(x, y, x_title = "X Axis", y_title = "Y Axis") {
df <- data.frame(x=q,y=p)

# Create the line plot
line_plot <- ggplot(df, aes(x = x, y = y)) + 
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  theme(
    axis.title.x = element_text(size = 20),  # Enlarge x-axis title
    axis.title.y = element_text(size = 20),  # Enlarge y-axis title
    axis.text.x = element_text(size = 25),   # Enlarge x-axis labels
    axis.text.y = element_text(size = 25)    # Enlarge y-axis labels
  )+
  labs(x = x_title, y = y_title)  # Set custom axis titles

# Create the summary table
summary_table <- data.frame(
  `P` = y,
 `Qd` = x
)

# Create the table using ggtexttable with larger text
table_plot <- ggtexttable(summary_table, rows = NULL, theme = ttheme("mOrange", base_size = 40))

# Arrange the line plot and table side by side
grid.arrange(line_plot, table_plot, ncol = 2, widths = c(3, 1))
}

# Call the function with custom x, y values and axis titles
x_values <- 1:10
y_values <- 2 * (1:10)
x_title = "Cups of Mocha per Month"
y_title = "Price in dollars"
draw_and_table(x_values, y_values, x_title , y_title  )
```

---

```{r, echo=F}
df <- data.frame(P=11:20, Q=1:10)
#create scatterplot with table
ggplot(df, aes(x=Q, y=P)) + 
    geom_point(aes(color=factor(P))) +
  geom_line()+
  # Custom y-axis labels
  theme(
    axis.title.x = element_text(size = 20),  # Enlarge x-axis title
    axis.title.y = element_text(size = 20),  # Enlarge y-axis title
    axis.text.x = element_text(size = 25),   # Enlarge x-axis labels
    axis.text.y = element_text(size = 25)) +   # Enlarge y-axis labels
  theme(legend.position = "none") +
    annotate(geom='table',
           x=12,
           y=20,
           label=list(df))
```

---
```{r, echo=F}

# Create the data frame
df <- data.frame(P = 11:20, Q = 1:10)

# Create the table using gridExtra with larger text
table <- tableGrob(df, rows = NULL, theme = ttheme_default(base_size = 20))  # Adjust base_size to make the table text bigger

# Create the scatterplot
p <- ggplot(df, aes(x = Q, y = P)) + 
  geom_point(aes(color = factor(P))) +
  geom_line() +
  xlim(1, 15) +  # Extend the x-axis to create space for the table
  ylim(10, 25) +  # Extend the y-axis to create space for the table
  theme(
    axis.title.x = element_text(size = 20),  # Enlarge x-axis title
    axis.title.y = element_text(size = 20),  # Enlarge y-axis title
    axis.text.x = element_text(size = 25),   # Enlarge x-axis labels
    axis.text.y = element_text(size = 25),   # Enlarge y-axis labels
    legend.position = "none"  # Remove the legend
  ) +
  annotation_custom(grob = table, xmin = 10, xmax = 14, ymin = 10, ymax = 20)  # Adjust positions to place the table

# Print the plot
print(p)
```
---
> "Teach a parrot the terms 'supply and demand' and you've got an economist."--- Thomas Carlyle


##  Lesson Objectives 

In this chapter, you will be able to answer the following:

+  What factors affect buyers' demand for goods?

+  What factors affect sellers' supply of goods? 

+  How do supply and demand determine the price of a good and the quantity sold?  

+  How do changes in the factors that affect demand or supply affect the market price and quantity of a good?  

+  How do markets allocate resources?

---
## DEMAND 


```{r , echo=FALSE,  fig.width=8.5, fig.height=6, warning=FALSE, message=F}
# Plot a user-defined function
myfun <- function(xvar) {
  xvar
}
# This sets the x range from 0 to 20
ggplot(data.frame(x = c(0, 20)), aes(x = x)) +
  stat_function(fun = myfun, geom = "line", linewidth = 1.5, linetype = "solid", color = "blue")+
  labs(x = "Quantity demanded, Q", y = "Price, P") +  # Custom axis titles
  scale_x_continuous(breaks = seq(0, 20, by = 5), labels = seq(0, 20, by = 5)) +  # Custom x-axis labels
  scale_y_continuous(breaks = seq(0, 20, by = 5), labels = seq(0, 20, by = 5))+  # Custom y-axis labels
  theme(
    axis.title.x = element_text(size = 20),  # Enlarge x-axis title
    axis.title.y = element_text(size = 20),  # Enlarge y-axis title
    axis.text.x = element_text(size = 25),   # Enlarge x-axis labels
    axis.text.y = element_text(size = 25))    # Enlarge y-axis labels
```

---
## Demand Schedule and Curve
```{r warning=FALSE, message=FALSE, echo=F}
# raw data vertical
 x = 1:7;
  y = 2*x
mydata <- data.frame(x,y)
# Transpose the data frame and create the table with column lines
as.data.frame(t(mydata)) %>%
  kbl(col.names = NULL, format = "html") %>%
  kable_styling(position = "center", full_width = FALSE) %>%
  column_spec(1:ncol(mydata), border_right = TRUE)  # Add column lines between cells
```


```{r , echo=FALSE}
# Plot a user-defined function
myfun <- function(xvar) {
  xvar
}
# This sets the x range from 0 to 20
ggplot(data.frame(x = c(0, 20)), aes(x = x)) +
  stat_function(fun = myfun, geom = "line")
```

---
## Demand analysis
.pull-left[


```{r , echo=FALSE,  fig.width=8.5, fig.height=6, warning=FALSE, message=F}
# Plot a user-defined function
myfun <- function(xvar) {
  xvar
}
# This sets the x range from 0 to 20
ggplot(data.frame(x = c(0, 20)), aes(x = x)) +
  stat_function(fun = myfun, geom = "line", linewidth = 1.5, linetype = "solid", color = "blue")+
  labs(x = "Quantity demanded, Q", y = "Price, P") +  # Custom axis titles
  scale_x_continuous(breaks = seq(0, 20, by = 5), labels = seq(0, 20, by = 5)) +  # Custom x-axis labels
  scale_y_continuous(breaks = seq(0, 20, by = 5), labels = seq(0, 20, by = 5))+  # Custom y-axis labels
  theme(
    axis.title.x = element_text(size = 20),  # Enlarge x-axis title
    axis.title.y = element_text(size = 20),  # Enlarge y-axis title
    axis.text.x = element_text(size = 25),   # Enlarge x-axis labels
    axis.text.y = element_text(size = 25))    # Enlarge y-axis labels
```


]

.pull-right[
```{r warning=FALSE, message=FALSE, echo=F}
# raw data vertical
 x = 1:7;
  y = 2*x
mydata <- data.frame(x,y)
as.data.frame(t(mydata)) %>% kable(col.names = NULL) %>% kable_styling()
```

]



```{r , echo=FALSE}
knitr::knit_exit()
```
